\textbf{Discuss the solutions of the Dirac equation for free particles and
their physical interpretation.}

\underline{Dirac equation}
\begin{itemize}
    \item SE derive for non-relativistic particles.
    \item Relaticistic energy: $E = p^2 c^2 + m^2 c^4$
    \item Klein-Gordon equation: 
    \begin{itemize}
        \item second order in time, negative probability densities.
        \item Negative energy solutions $\Rightarrow$ emmitting radiation without limit.
        \item Not stable in vacuum and overlooks spin.
    \end{itemize}
    \item Solution: Dirac equation 
    \begin{itemize}
        \item First order in time and space.
        \item Write energy $\vec{p}^2c^2 + m^2 c^4 = (c \alpha_x p_x + c \alpha_y p_y + c \alpha_z p_z + \beta mc^2)^2$
        \item Matrices $\alpha_i$ and $\beta$ satisfy anticommutation relations. Results in 
        \begin{align*}
        \alpha_i = \begin{pmatrix}
            0 & \sigma_i \\
            \sigma_i & 0
        \end{pmatrix}, \quad
            \beta = \begin{pmatrix}
                I & 0 \\
                0 & -I
            \end{pmatrix}
        \end{align*}
        \item Covariant form of eqn.: \((i\hbar \gamma^{\mu} \partial_{\mu} - mc) \psi = 0\)
        \item Introduces spinor 
    \end{itemize}
\end{itemize}


\underline{Free particle solution}
\begin{itemize}
    \item Plane wave solution: \(\psi = u(p) e^{\frac{i}{\hbar}(p \cdot x -Et)}\)
    \item Time-independent Dirac equation: \((c \vec{\alpha} \cdot \vec{p} + \beta mc^2) u(p) = E u(p)\)
    \item Zero momentum case $\Rightarrow$ eigenvalues \(E = \pm mc^2\)
    \item Four independent solutions:
    \begin{itemize}
        \item Two positive energy solutions (spin up and down)
        \item Two negative energy solutions (spin up and down)
    \end{itemize}
    \item General solution for arbitrary momentum:
    % \begin{align*}
    %     u^{(s)}(p) = \sqrt{\frac{E + mc^2}{2mc^2}} \begin{pmatrix}
    %         \chi^{(s)} \\
    %         \frac{\vec{\sigma} \cdot \vec{p} c}{E + mc^2} \chi^{(s)}
    %     \end{pmatrix}, \quad s = 1,2
    % \end{align*}
    % \item Negative energy solutions:
    % \begin{align*}
    %     v^{(s)}(p) = \sqrt{\frac{E + mc^2}{2mc^2}} \begin{pmatrix}
    %         \frac{\vec{\sigma} \cdot \vec{p} c}{E + mc^2} \chi^{(s)} \\
    %         \chi^{(s)}
    %     \end{pmatrix}, \quad s = 1,2
    % \end{align*}
\end{itemize}

\underline{Physical interpretation}

\begin{itemize}
    \item Vacuum instability problem: Dirac sea of negative energy states.
    \item Hole theory: absence of electron in negative energy state interpreted as positron.
    \item Predicts existence of antimatter
    \item Feynman: Antiparticles moves backwards in time, negative energy solutions reinterpreted.
    \item Dirac equation accounts for intrinsic spin of particles.
    \item Need for extra term in total angular momentum, since orbital angular momentum alone does not commute with Dirac Hamiltonian.
\end{itemize}


